---
title: 'Homework 3: Logistic Regression in R'
author: "Supratik Pochampally"
output:
  html_document:
    df_print: paged
---

This is an R script that will run logistic regression on a data set of acceptance probabilities for Indian graduate school. The original data set was uploaded by Mohan S Archarya, and can be found at https://www.kaggle.com/mohansacharya/graduate-admissions.

### Step 1

```{r}
# Load the data
df <- read.csv("Admission_Predict.csv")
# Take a look at the first few rows
head(df)
```

### Step 2

This data set is fairly unbalanced because out of the 400 total students in the data set, 365 have a Chance.of.Admit > 0.5. This means that we have a ratio of 91.25% Chance.of.Admit > 0.5 and 8.75% Chance.of.Admit < 0.5 ratio.

```{r}
# Make Research a factor
df$Research <- factor(df$Research)
# Get rid of the Serial No column
df <- df[2:9]
# Make a new column that is a binary factor based on if Chance.of.Admit > 0.5
df$Admit <- factor(ifelse(df$Chance.of.Admit > 0.5, 1, 0))
# Output column names
names(df)
# Output summary
summary(df)
```

### Step 3

The graphs below show that both high GRE scores and high TOEFL scores have a strong correlation with Admits, and that they are both good predictors.  

```{r, warning=FALSE}
# Create side by side graphs
par(mfrow=c(1, 2))
# GRE vs. Admit
plot(df$Admit, df$GRE.Score, main="GRE Score vs. Admit", xlab="Admit", ylab="GRE Score")
plot(df$Admit, df$TOEFL.Score, main="TOEFL Score vs. Admit", xlab="Admit", ylab="TOEFL Score")
```

### Step 4

```{r}
# Use seed to guarantee same split every run
set.seed(1234)
# Divide into train/test data
i <- sample(1:nrow(df), 0.75*nrow(df), replace=FALSE)
train <- df[i, ]
test <- df[-i, ]
```

### Step 5

The following warnings occurred:

Warning: glm.fit: algorithm did not converge

Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred

This occurred because the Admit was used as a predictor. Because this predictor holds the probability of admission rather than the classification of 0 or 1 for not admitted vs. admitted, the model failed to converge. 


```{r}
# Build a model predicting Admit from all predictors
glm0 <- glm(Admit~., data=train, family=binomial)
```

### Step 6

```{r}
# Build a model predicting Admit from all predictors excluding Chance.of.Admit
glm1 <- glm(Admit~.-Chance.of.Admit, data=train, family=binomial)
```

### Step 7

Looking at some of the probabilities and test predictions, we can conclude that the model was able to differentiate between low probabilities (44% and 8%) and higher probabilities.

The correlation value between the predicted probabilities and Chance.of.Admit is 0.648545, which indicates a neither low nor high correlation, but is slightly positive. This means that we are unsure about the correlation between the predictions and the Chance.of.Admit probabilities.  

```{r}
# Calculate the probabilities and display a few 
probs <- predict(glm1, newdata=test, type="response")
# Display few probabilities and test probs
tail(df$Chance.of.Admit, n=10)
tail(probs, n=10)
# Calculate the correlation between the predicted probabilities and Chance.of.Admit
cor(probs, test$Chance.of.Admit)
```

### Step 8
```{r}
# Calculate the predicted probabilities and display a few
pred <- ifelse(probs > 0.5, 1, 0)
# Make binary predictions and output a table
table(pred, test$Admit)
# Calculate and output the accuracy
acc1 <- mean(pred==test$Admit)
print(paste("glm1 accuracy = ", acc1))
```

### Step 9
```{r}
# Include the library ROCR
library(ROCR)
# Calculate the predictions for the ROCR graph
p <- predict(glm1, newdata=test, type="response")
pr <- prediction(p, test$Admit)
# Use the TPR measure for sensitivity and FPR measure for specificity
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
# Plot the ROCR graph
plot(prf)
# Calculate and print the AUC based on performance
auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
print(paste("glm1 auc = ", auc))
```

### Step 10

Both of the graphs show that both Admit and Research have an impact on the statement of purpose of each student.  

```{r}
# Create side by side graphs
par(mfrow=c(1, 2))
# SOP vs. Admit
plot(df$Admit, df$SOP, main="SOP vs. Admit", xlab="Admit", ylab="SOP")
# SOP vs. Research
plot(df$Research, df$SOP, main="SOP vs. Research", xlab="Research", ylab="SOP")
```